NMRlipids Databank: Overlay Databank of Lipid Membrane Simulations Arising from Open Collaboration

### 2022-11-02 main.tex
Update text and figure
Updates from Overleaf

%Up to three levels of \textbf{subheading} are permitted. Subheadings should not be numbered.
\subsection{Overlay structure of the NMRlipids databank}
The three layer structure of the NMRlipids databank, illustrated in Fig.~\ref{fig:overlay} a, enables an efficient way to share and upcycle large MD simulation trajectories.  Layer 1 contains raw simulation data that can locate in any publicly available repository or server offering long term stability and permanent links to the data, such as digital object identifiers. Layer 2 contains all the relevant information on the simulations, including links to the raw data in layer 1 and relevant metadata describing the systems. Layer 3 is the application layer containing outputs from the databank, such as graphical user interface (\url{http://www.databank.nmrlipids.fi/}) and results from analyses desribed in sections below.

@@

\begin{figure}
    \centering
    \includegraphics[width = 180mm]{Figures/overlay.pdf}
    \caption{a) Structure of an overlay databank. 
    %Raw data is available from a publicly accessible repository (layer 1).
    %The core of the databank contains information on the location of the raw data and indexing of molecular %compositions and simulation details described with universal naming conventions (layer 2).
    %The content of databank can be viewed and analysed using external programs that automatically read the information and data from layers 1 and 2 (layer 3). Examples of such applications are the GitHub repository generating the results presented in this work and the graphical access to the databank.
    More detailed structure of the layer 2 in the NMRlipids databank is illustrated in Fig.~\ref{DatabankStructure} in the SI.
    b) Distribution of the lengths of the trajectories, total number of trajectories and total lenght of the simulations in the NMRlipids databank.
    c) Distribution of lipids present in the trajectories in the NMRlipids databank. Lipids occuring in five or less simulations ('others') are listed in the right. 
    d) Currently available binary mixtures in the NMRlipids databank. 
    e) Distribution of force fields in the simulations in the NMRlipids databank.
    The figures and numbers are created on 9th of May 2022.}
    \label{fig:overlay}
\end{figure}

\subsection{Content of the databank}
Currently the databank is composed of approximately 500 trajectories with the total length of approximately 231 microseconds of which most are contributed for the previous publications from the NMRlipids open collaboration~\cite{botan15,catte16,antila19,bacle21}. The distribution of lipids, force fields, length of the trjajectories and available binary mixtures are shown in Fig.~\ref{fig:overlay}.

>>>>

%Up to three levels of \textbf{subheading} are permitted. Subheadings should not be numbered.
\subsection{Programmatic access to MD simulation data of membranes composed of most common biologically abundant lipids}
NMRlipids databank is a community driven databank containing atomistic MD simulations of biologically relevant lipid membranes emerging from the NMRlipids open collaboration~\cite{botan15,ollila16,catte16,antila19,bacle21}. Nearly thousand simulation trajectories with the total length approaching half a millisecond can be accessed programmatically using Python API or graphical user interface (\url{http://www.databank.nmrlipids.fi/}). NMRlipids databank is the only databank of MD simulations with open programmatic access to its content enabling fully flexible analyses of its content. Following the FAIR principles and NMRlipids project protocol, not only the content and computer programs related to the databank but also the whole construction process of the databank are open access~\cite{botan15}. The NMRlipids databank will be always open for data and other contributions.

The distribution of different lipid types in simulations in the NMRlipids databank (Fig.~\ref{fig:overlay}C) resembles the biological abundancies, the PC being the most common lipid type followed by PE, PS, PG, PI, and other lipids depending on organism and organelle~\cite{vanmeer08}. 
Currently the databank is composed of approximately 500 trajectories with the total length of approximately 231 microseconds of which most are contributed for the previous publications from the NMRlipids open collaboration~\cite{botan15,catte16,antila19,bacle21}. The distribution of lipids, force fields, length of the trjajectories and available binary mixtures are shown in Fig.~\ref{fig:overlay}.

The three layer structure of the NMRlipids databank, illustrated in Fig.~\ref{fig:overlay} a, enables an efficient way to share and upcycle large MD simulation trajectories.  Layer 1 contains raw simulation data that can locate in any publicly available repository or server offering long term stability and permanent links to the data, such as digital object identifiers. Layer 2 contains all the relevant information on the simulations, including links to the raw data in layer 1 and relevant metadata describing the systems. Layer 3 is the application layer containing outputs from the databank, such as graphical user interface (\url{http://www.databank.nmrlipids.fi/}) and results from analyses desribed in sections below.

@@

\begin{figure}
    \centering
    \includegraphics[width = 180mm]{Figures/overlay2.pdf}
    \caption{a) Structure of an overlay databank. 
    %Raw data is available from a publicly accessible repository (layer 1).
    %The core of the databank contains information on the location of the raw data and indexing of molecular %compositions and simulation details described with universal naming conventions (layer 2).
    %The content of databank can be viewed and analysed using external programs that automatically read the information and data from layers 1 and 2 (layer 3). Examples of such applications are the GitHub repository generating the results presented in this work and the graphical access to the databank.
    More detailed structure of the layer 2 in the NMRlipids databank is illustrated in Fig.~\ref{DatabankStructure} in the SI.
    b) Distribution of the lengths of the trajectories, total number of trajectories and total lenght of the simulations in the NMRlipids databank.
    c) Distribution of lipids present in the trajectories in the NMRlipids databank. Lipids occuring in five or less simulations ('others') are listed in the right. 
    d) Currently available binary mixtures in the NMRlipids databank. 
    e) Distribution of force fields in the simulations in the NMRlipids databank.
    The figures and numbers are created on 9th of May 2022.}
    \label{fig:overlay}
\end{figure}
